### 2017-02-22 pubs/demand_analysis/paper.tex

\newcount\isrr          % isrr=0 submitted version, isrr=1 dep res rep isrr=3 final PEPM '94
\def\theisrr{\the\isrr}
\isrr=3

\documentstyle[fleqn]{fpca93}
\newcommand{\progsize}{\normalsize}
\newcommand{\typescriptstyle}{\scriptscriptstyle}
\newcommand{\subtypesize}{\scriptsize}
%\newcommand{\sm}{$\mbox{\tt \ \ \ \ \ }\begin{array}{llllllll}}
%\newcommand{\fm}{\end{array}$}
\newcommand{\subtypespace}{\hspace{0.0pt}}
\newcommand{\subtypesubstyle}{\scriptstyle}             % scriptscriptstyle/scriptstyle

\newcommand{\com}{\newcommand}
\com{\rcom}{\renewcommand}

\input{../phd/commands}
\input{../phd/special}

% -------- A figure with a line at top and bottom, a caption and a label
% -------- Syntax: \fig{figoptions}{caption}{label}{ ..... }
\com{\fig}[4]{
\begin{figure}[#1]
\begin{center}
\begin{minipage}{\columnwidth}
\rule[1ex]{\textwidth}{1pt}
%\vspace{1ex}
#4
\caption{#2 \label{#3}}
\vspace{1ex}
\rule[1ex]{\textwidth}{1pt}
\end{minipage}
\end{center}
\end{figure}
}

\com{\backup}{\hspace{-0.12cm}}
\rcom{\espace}{\vspace{4pt}}
\com{\itt}[1]{\mbox{\it #1}}
%\com{\Ss}{\sf{S}}
\com{\Ps}{\sf{P}}

\com{\OUTC}{\itt{OUTC}}
\com{\SEL}{\itt{SEL}}
\com{\SPINE}{\itt{SPINE}}
\com{\LBR}{\itt{LBR}}

\sloppy

\pagestyle{empty}

\begin{document}

\thispagestyle{empty}

\title{PERs from Projections for Binding-time Analysis}

\author{Kei Davis\\Computing Science Department\\
        University of Glasgow\\
        Glasgow G12 8QQ, UK\\
        kei@dcs.glasgow.ac.uk.}

\date{}
\maketitle

\begin{abstract}

{\it First-order\/} projection-based binding-time analysis has proven
genuinely useful in partial evaluation \cite{Lau91a,Lau91c}.  There
have been three notable generalisations of projection-based analysis to
higher order.  The first lacked a formal basis \cite{Mog89}; the second
used structures strictly more general that projections, namely {\it
partial equivalence relations\/} (PERs) \cite{HS91}; the third involved
a complex construction that gave rise to impractically large abstract
domains \cite{Dav93b}.  This paper presents a technique free of
these shortcomings:  it is simple, entirely projection-based,
satisfies a formal correctness condition, and gives rise to reasonably
small abstract domains.  Though the technique is cast in terms of
projections, there is also an interpretation in terms of PERs.
The principal limitation of the technique is the restriction to
{\it monomorphic\/} typing.

\end{abstract}

\section{Introduction and Background}

We take as given that binding-time analysis is essential for good
partial evaluation, and we do not address the issue of annotating
programs according to the results of analysis.  Numerous binding-time
analysis techniques have been proposed and implemented; we greatly
narrow the field of discussion by restricting attention to those for
which there is a formally stated notion of correctness that the
technique has been proven to satisfy.

Analysis techniques can usually be identified as being based on either a
non-standard denotational semantics or a non-standard typing.  Examples
in the latter category include those of Gomard \cite{Go92}, Jensen
\cite{Jen92}, the Nielsons \cite{NN88}, Schmidt \cite{Sch88}, and
Henglein and Mossin \cite{HM94}.  Our
focus is on those techniques based on non-standard interpretation, in
particular, those using projections or partial equivalence relations
(PERs) as the basic abstract values.

A domain {\it projection\/} is a continuous idempotent function that
approximates the identity.  Launchbury \cite{Lau88} hit upon the idea
of using projections to encode degrees of staticness of data.  The
basic idea is that a projection maps to $\bot$ that part of a data
structure that is dynamic (possibly not determined), and acts as the
identity on that part which is static (definitely determined).
Examples are the identity $\ID$, the greatest projection, which
specifies that values are entirely static; the constant $\bot$ function
$\BOT$, the least projection, which specifies that values are entirely
dynamic; and projections $\FST$ and $\SND$ on product domains,
defined by
\[
\FST\ (x,y) = (x,\bot)\ ,\ \ \ \ \ \ \SND\ (x,y) = (\bot,y)\ ,
\]
specifying staticness in the first and second components of pairs, 
respectively.  The nominal goal of analysis is,
given function $f$ denoted by some programming-language expression, and
projection $\D$ encoding the staticness of the argument of $f$, to
determine $\G$ satisfying the {\it safety condition\/}
$\G\o\f\we\f\o\D$.  For example, taking $\G$ to be $\SND$ satisfies
$\G\o\itt{swap}\we\itt{swap}\o\FST$ for $\itt{swap}$ defined by
$\itt{swap}~(x,y)=(y,x)$.  Taking $\G$ to be $\BOT$ always satisfies
the safety condition but tells nothing; greater $\G$ is more
informative.  Launchbury \cite{Lau91a} showed that this safety
condition satisfies, and in a sense which he formalises, is equivalent
to the correctness condition for binding-time analysis in the general
framework of Jones \cite{Jon88}.  Using projection-based analysis,
Launchbury implemented both monomorphic and polymorphic versions of a
partial evaluator for a first-order language.  His work culminated in
the first strongly-typed partial evaluator \cite{Lau91c}---strong
evidence for the value of the projection-based approach.

There have been three notable attempts to generalise Launchbury's
techniques to higher order.  The first was Mogensen's generalisation of
the polymorphic technique \cite{Mog89}.  Though successfully
implemented, there is no formal statement of what it means for the
analysis to be correct; even if such a statement were made, proving
correctness would likely be difficult because of the highly intensional
nature of the analysis:  the non-standard values associated with
expressions are strongly dependent on their syntactic structure, and
projections are encoded symbolically as {\it abstract closures\/}, with
approximation performed algebraically `on-the-fly' according to time
and space considerations.  Nonetheless, the experiment provided
evidence for the {\it practicality\/} of the projection-based approach
at higher order.

The second generalisation was Hunt and Sands', of the monomorphic
technique to higher order \cite{HS91}.  Their observation was that a
projection, regarded as a set of domain-range pairs, is an {\it
equivalence relation\/}:  given $\G$, values $u$ and $v$ are in the
same equivalence class if $\G~u=\G~v$, and the canonical elements of
the equivalence classes are the set of fixed points (range) of $\G$.
Hunt showed that the safety condition $\G\o\f\we\f\o\D$ holds iff $f$
is related to itself by $\D\fto\G$ where $\fto$ is the standard
operation on binary relations, so
\beqs
\it &\it (\G\o\f\we\f\o\D)\\
\it \Leftrightarrow&\it (\forall\ u,v\ .\ \D\ u=\D\ v\ \Rightarrow\ \G\ (f\ u)=\G\ (f\ v))\ .
\eeqs
Then, for example, $(\BOT\fto\ID)~(f,f)$ asserts that $f$ maps dynamic
arguments to static results.  In general $\D\fto\G$ is not an
equivalence relation, but it is always a {\it partial\/} equivalence
relation: it is symmetric and transitive but not necessarily
reflexive.  Unlike projections regarded as relations, PERs are closed
under $\fto$; the result, as Hunt and Sands show, is that `scaling up'
to higher-order analysis is reasonably straightforward.  One
disadvantage of their method is that PER spaces are considerably larger
than the projection spaces on the same domains, and it is not clear
which PERs to choose for (finite) abstract domains.  Here they borrowed
heavily from the projection world, using standard abstract projection
domains at ground types.  Further, its practicality has not been
demonstrated by implementation, and, because of the unfamiliar
territory, a promising route to a polymorphic generalisation is obscure.

The third generalisation was ours, to an entirely projection-based,
monomorphic, higher-order technique \cite{Dav93b}.  One observation
motivating the approach is that there is no meaningful abstraction of
values to projections, only of functions $f$ to projection transformers
$\T$ (functions from projections to projections) satisfying
$(\T~\D)\o\f\we\f\o\D$ for all $\D$.  To make this abstraction possible
a semantics intermediate between the standard and analysis semantics
was introduced.  Moving from the standard to intermediate semantics
involved a translation of each ground type \mbox{\tt T} to a function type
\mbox{\tt E}\ \mbox{\tt ->}\ \mbox{\tt T} (for a fixed type \mbox{\tt E}), the values of which, being
functions, could then be abstracted.  The result, while proven correct
with respect to a formal safety condition, is probably not practicable
because of the growth in the sizes of (usefully rich) abstract domains
induced by the type translation.

This paper presents a technique far simpler technique than our previous
one.  No intermediate semantics is required, and the correctness
condition and proof are much simpler.  Because the translation of
ground types to function types is avoided the abstract domains are much
smaller. Though entirely projection-based, we show that there is a
reading of the results in terms of PERs, intimating a close
relationship with the PER-based technique.

\section{Language and Standard Semantics}
The source language is a simple, strongly typed, monomorphic,
non-strict functional language.  The grammar for the language
of types and type definitions is given in Figure~\ref{fig:types}.
\fig{tb}{Types and type definitions.}{fig:types}{
\beqs
\it \makebox{\tt T}&\it ::=&\it \ta&\it \reason{Type\ Name}\\
\it &\it |&\it \tint&\it \reason{Integer}\\
\it &\it |&\it \tprod&\it \reason{Product,\ $n\geq{\rm0}$}\\
\it &\it |&\it \tsum&\it \reason{Sum,\ $n\geq{\rm1}$}\\
\it &\it |&\it \tfun&\it \reason{Function}\ \espace\\
\it \makebox{\tt D}&\it ::=&\it \tdefn&\it \reason{Type\ Definitions}
\eeqs
}
Nullary product corresponds to the so-called {\it unit\/} type.  A unary
product \mbox{\tt (T)} will always have the same semantics as \mbox{\tt T}.
The types used in the examples are defined as follows.
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ FunList\ =\ \ \ nil\ ()\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ +\ cons\ (Int\ ->\ Int,\ FunList)\ ,\ }\vspace{-0.0em}\\
\makebox{\progsize\tt }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ FunTree\ =\ \ \ leaf\ (Int\ ->\ Int)\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ +\ branch\ (FunTree,\ FunTree)\ .}\vspace{-0.0em}
\end{flushleft}

The grammar for expressions is given in Figure~\ref{fig:exprs}.
Addition for integers is provided as typical of operations on flat data
types in this setting.  A unary tuple \mbox{\tt (e)} will always have the same
semantics as \mbox{\tt e}.  The (monomorphic) typing of expressions is entirely
standard and is omitted.

\fig{tb}{Expressions.}{fig:exprs}{
\beqs
\it \makebox{\tt e}&\it ::=&\it \evar&\it \reason{Variable}\\
\it &\it |&\it \enum&\it \reason{Numeral}\\
\it &\it |&\it \esum&\it \reason{Integer\ addition}\\
\it &\it |&\it \eapp&\it \reason{Function\ application}\\
\it &\it |&\it \etup&\it \reason{Tuple\ construction}\\
\it &\it |&\it \makebox{\tt let}\ \makebox{\tt (}\xone\makebox{\tt ,}\ldots\makebox{\tt ,}\xn\makebox{\tt )}\ \makebox{\tt =}\ \enot\\
\it &\it &\it \makebox{\tt \ \ in}\ \eone&\it \reason{Tuple\ decomposition}\\
\it &\it |&\it \econ&\it \reason{Sum\ construction}\\
\it &\it |&\it \makebox{\tt case}\ \enot\ \makebox{\tt of}\\
\it &\it &\it \makebox{\tt \ \ }\{\ci\ \xxi\ \makebox{\tt ->}\ \ei\}&\it \reason{Sum\ decomposition}\\
\it &\it |&\it \elamt&\it \reason{Lambda\ abstraction}\\
\it &\it |&\it \eapp&\it \reason{Function\ application}\\
\it &\it |&\it \efix&\it \reason{Fixed\ point}
\eeqs
}

\subsection{Expression semantics}

Since two different expression semantics will be given, following
Abramsky \cite{Abr90} we define a semantics $\cE{}$ parameterised by a
set of {\it defining constants}.  The semantics $\cE{}$ is defined in
Figure~\ref{fig:gensem}; the defining constants are {\it
plus}, $\itt{sel}_i$, {\it tuple}, $\itt{inc}_i$, $\itt{outc}_i$, {\it
choose}, {\it mkfun}, {\it apply}, and {\it fix}.  The two instances of
$\cE{}$ are distinguished by a superscript:  $\cE{\Ss}$ for the
standard semantics and $\cE{\Ps}$ for the non-standard semantics.  The
corresponding type semantics have the same superscripts, as do the
defining constants.

It is useful to regard the free-variable environment of each expression
as having some tuple type $\tprod$, and environment lookup as indexing
(as in a categorical semantics, or De~Bruijn indexing); variables are
indexed implicitly or explicitly by their index in the free variable
environment.  Then for both versions of the semantics and all
expressions \mbox{\tt e} of type \mbox{\tt T} with environment of type $\tprod$,
\beqs
\it \Ee{}{\makebox{\tt e}}\ \in\ \Te{}{\tprod}~\fto~\Te{}{\makebox{\tt T}}\ .
\eeqs
Noting that $\R\br{\xxi}$ is short for $\itt{sel}_i~\R$, environment
update $\R[\xxi\mapsto{}v]$ is defined by
\beqqs
\it tuple\ (&\it sel_{\rm1}\ \R,\ \ldots,\ sel_{i-{\rm1}}\ \R,\\
\it &\it v,\\
\it &\it sel_{i+{\rm1}}\ \R,\ \ldots,\ sel_n\ \R)\ .
\eeqqs
It is convenient to regard \mbox{\tt Int} as being defined as the infinite sum
\beqs
\it \tint\makebox{\tt \ =\ }\ldots\makebox{\tt \ +\ n}_{-{\rm1}}\makebox{\tt \ ()\ +\ n}_{{\rm0}}\makebox{\tt \ ()\ +\ n}_{{\rm1}}\makebox{\tt \ ()\ +\ }\ldots\ ,
\eeqs
where in practice we write $\mbox{\tt n}_i$ as short for $\mbox{\tt n}_i~\mbox{\tt ()}$.

\fig{tb}{Parameterised semantics.}{fig:gensem}{
\beqs
\it \Ee{}{\evar_i}\ \R\ =\ \R\br{\makebox{\tt x}_i}\ =\ \itt{sel}_i\ \R\espace\espace\\
\it %\ \Ee{}{\eunit}\ \R\ =\ \itt{mkunit}\ \R\espace\espace\\
\it %\ \Ee{}{\enum}\ \R\ =\ \itt{mkint}_i\espace\espace\\
\it \Ee{}{\esum}\ \R\ =\ \itt{plus}~(\Ee{}{\eone}\ \R,\ \Ee{}{\etwo}\ \R)\espace\espace\\
\it \Ee{}{\etup}\ \R\\
\it \ \ \ \ \ =\ \itt{tuple}~(\Ee{}{\eone}\ \R,~\ldots,~\Ee{}{\en}\ \R)\espace\espace\\
\it \Ee{}{\esel}\ \R\\
\it \ \ \ \ \ =\ \Ee{}{\eone}\ \R[\xxi\mapsto\itt{sel}_i\ (\Ee{}{\enot}\ \R)~|~\oin]\espace\espace\\
\it \Ee{}{\econ}\ \R\ =\ \itt{inc}_i\ (\Ee{}{\makebox{\tt e}}\ \R)\espace\espace\\
\it \Ee{}{\ecase}\ \R\\
\it \ \ \ \ \ \ =\ \itt{choose}~(\ba[t]{l}\Ee{}{\enot}\ \R,\\
\it \Ee{}{\eone}\ \R[\xone\mapsto\itt{outc}_{\rm1}\ (\Ee{}{\enot}\ \R)],\\
\it \ \ \ ~\vdots\\
\it \Ee{}{\en}\ \R[\xn\mapsto\itt{outc}_n\ (\Ee{}{\enot}\ \R)])\ea\espace\espace\\
\it \Ee{}{\elam}\ \R\ =\ \itt{mkfun}\ (\L{}x.\Ee{}{\makebox{\tt e}}\ \R[\evar\mapsto{}x])\espace\espace\\
\it \Ee{}{\eapp}\ \R\ =\ \itt{apply}\ (\Ee{}{\eone}\ \R)\ (\Ee{}{\etwo}\ \R)\espace\espace\\
\it \Ee{}{\efix}\ \R\ =\ (\itt{fix}\o\itt{apply})\ (\Ee{}{\makebox{\tt e}}\ \R)
\eeqs
}

\subsection{Standard semantics}

The standard $\Ss$ type and expression semantics are defined in
Figure~\ref{fig:ssem}.  Function types give rise to lifted function
spaces as in Abramsky's lazy lambda calculus \cite{Abr89}, and the
semantics distinguishes those expressions of function type that have
WHNF (have value $\itt{lift}~f$ for some $f$) and those that do not (have
value $\bot$).  Products are unlifted; a unary sum-of-products gives a
lifted product.  Domain $\plus$ is separated sum.  Recursive type
definitions give rise to recursive domain specifications which have the
usual least-fixed-point solutions.

\fig{tb}{Standard type and expression semantics.}{fig:ssem}{
\beqs
\it \Te{\Ss}{\tint}\ =\ \Int\ =\ \lift{\bf{Z}}\ \espace\espace\\
\it \Te{\Ss}{\tprod}\ =\ \Te{\Ss}{\tone}\ \times\ \ldots\ \times\ \Te{\Ss}{\tn}\espace\espace\\
\it \Te{\Ss}{\tsum}\\
\it \ \ \ \ \ =\ \Te{\Ss}{\tone}\ \plus\ \ldots\ \plus\ \Te{\Ss}{\tn}\espace\espace\\
\it \Te{\Ss}{\tfun}\ =\ \liftp{\Te{\Ss}{\tone}\ \fto\ \Te{\Ss}{\ttwo}}\\
\it \\
\it %\ \itt{mkunit}^{\Ss}\ \R\ =\ ()\espace\espace\\
\it %\ \itt{mkint}^{\Ss}_n\ \R\ =\ \itt{lift}\ n\espace\espace\\
\it \itt{plus}^{\Ss}\ (x,y)\ =\ x+y\espace\espace\\
\it \itt{tuple}^{\Ss}\ (x_{\rm1},\ldots,x_n)\ =\ (x_{\rm1},\ldots,x_n)\espace\espace\\
\it \itt{sel}_i^{\,\Ss}\ (x_{\rm1},\ldots,x_n)\ =\ x_i\espace\espace\\
\it \itt{inc}_i^{\Ss}\ =\ \itt{in}_i\o\itt{lift}\espace\espace\\
\it \itt{outc}_i^{\Ss}\ x\ =\ \itt{drop}\o\itt{out}_i\espace\espace\\
\it \ba{lll}\itt{choose}^{\Ss}\ (\bot,&\it \ x_{\rm1},\ldots,x_n)\ =\ \bot\\
\it \itt{choose}^{\Ss}\ (\itt{in}_i~v,&\it \ x_{\rm1},\ldots,x_n)\ =\ x_i\ea\espace\espace\\
\it \itt{mkfun}^{\Ss}\ =\ \itt{lift}\espace\espace\\
\it \itt{apply}^{\Ss}\ =\ \itt{drop}\espace\espace\\
\it \itt{fix}^{\Ss}\ =\ \itt{lfp}\ \reason{Least\ fixed\ point}
\eeqs
}

\section{Domain Factorisation}

A key observation of \cite{Dav93b} was that since there is no concept
of staticness of the body of a lambda expression, there is no point in
having projections on function spaces.  Hence domains are factored into
their evaluable or {\it data\/} parts, and their unevaluable but
applicable {\it forward\/} parts.  For example, for
$\liftp{T\fto{}U}=\Te{\Ss}{\mbox{\tt T}\ \mbox{\tt ->}\ \mbox{\tt U}}$ we need only distinguish two
degrees of definedness---between $\bot$ and values of the form
$\itt{lift}~f$.  This may be encoded by $\lone$---the data part of
$\liftp{T\fto{}U}$---on which there are precisely two projections,
namely $\ID$ and $\BOT$.  Here the forward part is $T\fto{}U$.

The {\it data domain\/} corresponding to type \mbox{\tt T} is $\De{}{\mbox{\tt T}}$,
where $\cD{}$ is defined exactly like $\cT{\Ss}$ except that function
spaces are replaced by the one-point domain $\one=\{\unit\}$, that is,
$\De{}{\tfun}=\lift{\one}$.  The function $\data$ from values in
$\Te{\Ss}{\mbox{\tt T}}$ to their data parts in $\De{}{\mbox{\tt T}}$ is a projection:
it is like the identity except that values from function spaces are
mapped into $\one$.  The projection $\data$ is defined in terms of the
structure of \mbox{\tt T} as follows.
\beqs
\it {\it{}data}_{\subtypespace\stint}\ =\ {\it{}id}_{\subtypespace\stint}\ ,\espace\\
\it {\it{}data}_{\subtypespace\stprod}\ =\ {\it{}data}_{\subtypespace\stone}\ \times\ \ldots\ \times\ {\it{}data}_{\subtypespace\stn}\ ,\espace\\
\it {\it{}data}_{\subtypespace\stsum}\ =\ {\it{}data}_{\subtypespace\stone}\ \plus\ \ldots\ \plus\ {\it{}data}_{\subtypespace\stn}\ ,\espace\\
\it {\it{}data}_{\subtypespace\stfun}\ =\ {\liftp{\Lx.\unit}}\ .
\eeqs
The last definition uses function lifting, defined by $\lf~\bot=\bot$
and $\lf~(\itt{lift}~x)=\itt{lift}~(f~x)$.  Recursive type definitions
give rise to recursive function specifications with only one solution.

In the same style as  $\De{}{\mbox{\tt T}}$ and $\data$ we define
$\Tde{\Ss}{\mbox{\tt T}}$ to give the forward domain for \mbox{\tt T}, and $\fun$ to be
the function mapping values in $\Te{\Ss}{\mbox{\tt T}}$ to their forward parts
in $\Tde{\Ss}{\mbox{\tt T}}$, as follows.  Roughly, $\cTd{\Ss}$ is like
$\cT{\Ss}$ with all lifting removed and sum replaced by product.
\beqs
\it \Tde{\Ss}{\tint}\ =\ \one\ ,\espace\\
\it \Tde{\Ss}{\tprod}\ =\ \Tde{\Ss}{\tone}\ \times\ \ldots\ \times\ \Tde{\Ss}{\tn}\ ,\espace\\
\it \Tde{\Ss}{\tsum}\\
\it \ \ \ \ \ =\ \Tde{\Ss}{\tone}\ \times\ \ldots\ \times\ \Tde{\Ss}{\tn}\ ,\espace\\
\it \Tde{\Ss}{\tfun}\ =\ \Te{\Ss}{\tone}\ \fto\ \Te{\Ss}{\ttwo}\ .
\eeqs
The mapping from standard values to their forward parts is defined by
\beqs
\it {\it{}fun}_{\subtypespace\stint}\ =\ \Lx.\unit\ ,\espace\\
\it {\it{}fun}_{\subtypespace\stprod}\ =\ {\it{}fun}_{\subtypespace\stone}\ \times\ \ldots\ \times\ {\it{}fun}_{\subtypespace\stn}\ ,\espace\\
\it \ba{lll}{\it{}fun}_{\subtypespace\stsum}\ \bot&\it =\ \bot\espace\\
\it {\it{}fun}_{\subtypespace\stsum}\ (\itt{in}_i~v)\\
\it \ \ \ \ \ =\ (\bot,\ldots,\bot,v,\bot_,\ldots,\bot)&\it \reason{$v$\ in\ $\ith$\ position}\ ,\ea\espace\\
\it {\it{}fun}_{\subtypespace\stfun}\ =\ \itt{drop}\ .
\eeqs
We write $\fac$ for $\Lx.(\data~x,\ \fun~x)$.  Then
$\De{}{\mbox{\tt T}}\times\Tde{\Ss}{\mbox{\tt T}}$ is a factorisation of $\Te{\Ss}{\mbox{\tt T}}$,
and
\beqs
\it \fac\ \in\ \Te{\Ss}{\makebox{\tt T}}\ \fto\ (\De{}{\makebox{\tt T}}\ \times\ \Tde{\Ss}{\makebox{\tt T}})
\eeqs
is an embedding which {\it determines\/} the corresponding projection
$\unfac$ (since they are related by $\unfac\o\fac=\itt{id}$ and
$\fac\o\unfac\we\itt{id}$).  Rather than give an explicit definition of
$\unfac$ we give some examples.  The factorisation of $\Int$ is
$\Int\times\one$; more generally, the factorisation of any domain $D$
corresponding to a type not containing \mbox{\tt ->} is just $D\times\one$.  If
$T=\Te{\Ss}{\mbox{\tt T}}$ and $U=\Te{\Ss}{\mbox{\tt U}}$ then the factorisation of
$\lift{(T\fto{}U)}=\Te{\Ss}{\mbox{\tt T}\ \mbox{\tt ->}\ \mbox{\tt U}}$ is
$\lift{\one}\times(T\fto{}U)$---note that factorisation `stops' at
function-space constructors.  Now
${\it{}fac}_{\subtypespace{\makebox{\subtypesize\tt T->U}} }~\bot=(\bot,\bot)$, and
${\it{}fac}_{\subtypespace{\makebox{\subtypesize\tt T->U}} }~(\itt{lift}~f)=(\lunit,~f)$ for all $f$.  In
the other direction ${\it{}unfac}_{\subtypespace{\makebox{\subtypesize\tt T->U}} }~(\bot,f)=\bot$ and
${\it{}unfac}_{\subtypespace{\makebox{\subtypesize\tt T->U}} }~(\lunit,~f)=(\itt{lift}~f)$; this must be
since in general $\unfac\o\fac$ is the identity.

\section{Projection Semantics}

Standard values $v$ and $v'$ are related by nonstandard value $(\A,\K)$
when their data parts are related by $\A$, and their forward parts are
logically related by $\K$.  At each type \mbox{\tt T} the relation is
$\Re{}{\mbox{\tt T}}~((\A,\K),\cdot,\cdot)$, defined by
\beqs
\it \lefteqn{\Re{}{\makebox{\tt T}}~((\A,\K),v,v')}\\
\it \ \ \ \ \ =&\it \mit(\A~d=\A~d')\ \wedge\ \Rde{}{\makebox{\tt T}}~(\K,f,f')\\
\it &\it \mbox{\rm{}where}\\
\it &\it \ \ \ \ba[t]{lll}(d,f)&\it \:=\ \fac~v\\
\it \mit(d',f')&\it \:=\ \fac~v'\ ,\ea
\eeqs
where $\Rde{}{\mbox{\tt T}}$ is defined by
\beqs
\it \Rde{}{\tint}~((),(),())\ =\ \itt{True}\ ,\espace\\
\it \Rde{}{\tprod}\ =\ \Rde{}{\tone}\ \x\ \ldots\ \x\ \Rde{}{\tn}\ ,\espace\\
\it \Rde{}{\tsum}\\
\it \ \ \ \ \ =\ \Rde{}{\tone}\ \x\ \ldots\ \x\ \Rde{}{\tn}\ ,\espace\\
\it \Rde{}{\tfun}\ =\ \Re{}{\tone}\ \fto\ \Re{}{\ttwo}\ .
\eeqs
Here $\x$ and $\fto$ are the standard operations on (ternary)
relations:  for relations $R$ and $S$ we have
$(R\times{}S)((x,y),(x',y'),(x'',y''))$ iff $R(x,x',x'')$ and
$S(y,y',y'')$, and $(R\fto{}S)(f,g,h)$ iff for all $x$, $y$, and $z$
such that $R(x,y,z)$ we have $S(f~x,\ g\ y,\ h~z)$.  Here, recursive
type definitions give recursive relation specifications, which have
inclusive least-fixed-point solutions. (A relation is inclusive if,
when it holds for each element of an ascending chain, it also holds at
the limit.  Such relations are sometimes called {\it admissible\/} or
{\it chain complete}.  Some work is required to show inclusivity of
these recursively-defined relations.) Thus the $\Ps$ type semantics
$\cT{\Ps}$ must be
\beqs
\it \Te{\Ps}{\makebox{\tt T}}\ =\ \proj{\De{}{\makebox{\tt T}}}\ \times\ \Tde{\Ps}{\makebox{\tt T}}\ ,
\eeqs
where $\proj{\De{}{\mbox{\tt T}}}$ is the lattice of projections on domain
$\De{}{\mbox{\tt T}}$, and $\cTd{\Ps}$ is defined exactly like $\cTd{\Ss}$ with
superscript $\Ps$ everywhere replacing superscript $\Ss$.

\fig{tb}{Projection semantics.}{fig:fsem}{
\beqs
\it %\ \itt{mkunit}^{\Ps}\ \R\ =\ ()\\
\it %\espace\espace\\
\it %\ \itt{mkint}^{\Ps}_n\ \R\ =\ \itt{lift}\ n\\
\it %\espace\espace\\
\it \itt{plus}^{\Ps}\ ((\A,()),(\B,()))\\
\it \ \ \ =\ \left\{\ba{ll}(\ID,()),&\it \ \ \ \ \mbox{\rm\ if\ $\A=\ID$\ and\ $\B=\ID$}\\
\it (\BOT,()),&\it \ \ \ \ \mbox{\rm\ otherwise}\ea\right.\espace\\
\it \itt{tuple}^{\Ps}\ ((\A_{\rm1},\K_{\rm1}),\ldots,(\A_n,\K_n))\\
\it \ \ \ =\ ((\A_{\rm1}\x\ldots\x\A_n),(\K_{\rm1},\ldots,\K_n))\espace\\
\it \itt{sel}_i^{\,\Ps}\ ((\A_{\rm1}\x\ldots\x\A_n),(\K_{\rm1},\ldots,\K_n))\ =\ (\A_i,\K_i)\espace\\
\it \itt{inc}_i^{\Ps}\ (\A,\K)\ =\ (C_i\ \A,\ (\top,\ldots,\top,\K,\top,\ldots,\top))\espace\\
\it \itt{outc}_i^{\Ps}\ (\A,(\K_{\rm1},\ldots,\K_n))\ =\ (\OUTC_i\ \A,\ \K_i)\espace\\
\it \itt{choose}^{\Ps}\ ((\A,\K),x_{\rm1},\ldots,x_n)\\
\it \ \ \ \ \ =\ \left\{\ba{ll}\bot,&\it \ \ \ \ \mbox{\rm\ if}\ \A\not\se\glb_{\oin}\ C_i~\BOT\\
\it x_{\rm1}\glb\ldots\glb{}x_n,&\it \ \ \ \ \mbox{\rm\ otherwise}\ea\right.\espace\\
\it \itt{mkfun}^{\Ps}\ f\ =\ (\ID,f)\espace\\
\it \itt{apply}^{\Ps}\ (\A,f)\ =\ \left\{\ba{lll}\bot,&\it \ \ \ \ \mbox{\rm\ if}\ \A=\BOT\\
\it f,&\it \ \ \ \ \mbox{\rm\ if}\ \A=\ID\ \ea\right.\espace\\
\it \itt{fix}^{\Ps}\ =\ \itt{gfp}\ \reason{Greatest\ fixed\ point}
\eeqs
}

The defining constants for the projection semantics $\cE{\Ps}$ are
given in Figure~\ref{fig:fsem}. The following notation is used for
specifying projections on the data domains.  For sum type $\tsum$ with
data domain $D_1\plus\ldots\plus D_n$ define $C_i~\A$ to be
$\ID+\cdots+\ID+\A+\ID+\cdots+\ID$ where $\A$ is the $i^{th}$ summand,
$\OUTC_i~(\G_1+\cdots+\G_n)$ to be $\G_i$, and $\OUTC_i~\BOT$ to be
$\BOT$.

\ \\
\noindent {\bf The Central Result.}  For expression \mbox{\tt e} of type \mbox{\tt T} with free-variable
environment of type \mbox{\tt E}, the functions $\Ee{\Ps}{\mbox{\tt e}}$,
$\Ee{\Ss}{\mbox{\tt e}}$, and $\Ee{\Ss}{\mbox{\tt e}}$ are logically related by $\cR{}$,
that is
\beqs
\it (\Re{}{\makebox{\tt E}}\fto\Re{}{\makebox{\tt T}})\ (\Ee{\Ps}{\makebox{\tt e}},\ \Ee{\Ss}{\makebox{\tt e}},\ \Ee{\Ss}{\makebox{\tt e}})\ .
\eeqs
Further, $\Re{}{\mbox{\tt T}}~((\A,\K),\cdot,\cdot)$ is a PER for all $\A$ and $\K$.\ \ $\Box$

\ \\
\noindent The bulk of proof is omitted; it consists of showing that the defining
constants are similarly related, and a simple induction on the
structure of expressions showing that if the defining constants are
logically related, then so are the semantic functions.  By way of example
we consider the expression form $\esum$, and the relevant defining constant
$\it plus$.
To show that $\it plus^{\Ss}$ and $\it plus^{\Ps}$ are correctly related
we need to show that
\beqqs
\it ((\Re{}{\tint}\x\Re{}{\tint})\fto\Re{}{\tint})\ (&\it plus^{\Ps},\\
\it &\it plus^{\Ss},\\
\it &\it plus^{\Ss})\ ,
\eeqqs
that is, for all $\A_1$, $\A_2$, $d_1$, $d_2$, $d_1'$, and $d_2'$ that
\beqs
\it &\it \mit(\ID\ d_{\rm1}=\ID\ d_{\rm1}')\ \land\ (\ID\ d_{\rm2}=\ID\ d_{\rm2}')\\
\it \limp&\it \mit\ID\ (d_{\rm1}+d_{\rm2})=\ID\ (d_{\rm1}'+d_{\rm2}')
\eeqs
and for $\A_1\neq\ID$ or $\A_2\neq\ID$ that
\beqs
\it &\it \mit(\A_{\rm1}\ d_{\rm1}=\A_{\rm1}\ d_{\rm1}')\ \land\ (\A_{\rm2}\ d_{\rm2}=\A_{\rm2}\ d_{\rm2}')\\
\it \limp&\it \mit\BOT\ (d_{\rm1}+d_{\rm2})=\BOT\ (d_{\rm1}'+d_{\rm2}')\ .
\eeqs
For the inductive case $\esum$ with environment type \mbox{\tt E} we need to show that if
\beqs
\it (\Re{}{\makebox{\tt E}}\fto\Re{}{\tint})\ (\Ee{\Ps}{\eone},\ \Ee{\Ss}{\eone},\ \Ee{\Ss}{\eone})
\eeqs
and
\beqs
\it (\Re{}{\makebox{\tt E}}\fto\Re{}{\tint})\ (\Ee{\Ps}{\etwo},\ \Ee{\Ss}{\etwo},\ \Ee{\Ss}{\etwo})
\eeqs
then
\beqqs
\it (\Re{}{\makebox{\tt E}}\fto\Re{}{\tint})\ (&\it \Ee{\Ps}{\esum},\\
\it &\it \Ee{\Ss}{\esum},\\
\it &\it \Ee{\Ss}{\esum})\ .
\eeqqs
Suppose $\Re{}{\mbox{\tt E}}~(\R^{\Ps},\R^{\Ss},\R'^{\Ss})$.  By the induction
hypothesis we have
$\Re{}{\tint}~(\A_i,v_i,v'_i)$, where 
$(\A_i,())=\Ee{\Ps}{\ei}~\R^{\Ps}$,
$(v_i,())=\Ee{\Ps}{\ei}~\R^{\Ss}$, and
$(v'_i,())=\Ee{\Ps}{\ei}~\R'^{\Ss}$, for $i=1,2$.
Now
\beqqs
\it \Re{}{\tint}\ (&\it \Ee{\Ps}{\esum}\ \R^{\Ps},\\
\it &\it \Ee{\Ss}{\esum}~\R^{\Ss},\\
\it &\it \Ee{\Ss}{\esum}~\R'^{\Ss})
\eeqqs
iff
\beqqs
\it \Re{}{\tint}\ (&\it plus^{\Ps}\ ((\A_{\rm1},()),(\A_{\rm2},())),\\
\it &\it plus^{\Ss}\ \mit((v_{\rm1},()),(v_{\rm2},())),\\
\it &\it plus^{\Ss}\ \mit((v'_{\rm1},()),(v'_{\rm2},())))\ ,
\eeqqs
which holds since $\it plus^{\Ps}$ and $\it plus^{\Ps}$, and their various
arguments, are correctly related.

\section{Abstract Domains}

At each type \mbox{\tt T} we require a finite abstraction of $\Te{\Ps}{\mbox{\tt T}}$.
This abstract domain is ${\it{}FProj}_{\subtypespace\stt}\times{\it{}FFor}_{\subtypespace\stt}$,
where ${\it{}FProj}_{\subtypespace\stt}$ is a finite abstraction of the lattice
$\proj{\De{}{\mbox{\tt T}}}$, and ${\it{}FFor}_{\subtypespace\stt}$ is a finite abstraction
of $\Tde{\Ps}{\mbox{\tt T}}$.  The definition of ${\it{}FProj}_{\subtypespace\stt}$ is based
on that in \cite{Lau91a}.  A projection $\G$ is in
${\it{}FProj}_{\subtypespace\stt}$ if $\G~\mbox{\bf proj}~\mbox{\tt T}$ can be inferred from
the rules given in Figure~\ref{fig:infrules}.
%
\fig{tb}{Inference rules for finite projection domains.}{fig:infrules}{
\beqs
\it \begin{array}{lll}\axm{\BOT\ \ \mbox{\bf{}proj}\ \ \tint}\ \ \ \ \ \ \ \ \ \ &\it \axm{\ID\ \ \mbox{\bf{}proj}\ \ \tint}\espace\espace\\
\it \axm{\BOT\ \ \mbox{\bf{}proj}\ \ \tfun}&\it \axm{\ID\ \ \mbox{\bf{}proj}\ \ \tfun}\end{array}\espace\espace\\
\it \axm{\BOT\ \ \mbox{\bf{}proj}\ \ \tsum}\espace\espace\\
\it \infero{\G_{\rm1}\ \ \mbox{\bf{}proj}\ \ \tone\ \ \cdots\ \ \G_n\ \ \mbox{\bf{}proj}\ \ \tn}{(C_{\rm1}\ \G_{\rm1})\glb\cdots\glb(C_n\ \G_n)\ \ \mbox{\bf{}proj}\ \ \tsum}\espace\espace\\
\it \infero{\G_{\rm1}\ \ \mbox{\bf{}proj}\ \ \tone\ \ \cdots\ \ \G_n\ \ \mbox{\bf{}proj}\ \ \tn}{\G_{\rm1}\x\ldots\x\G_n\ \ \mbox{\bf{}proj}\ \ \tprod}\espace\espace\\
\it \infer{\ba{ll}&\it \G_{\rm1}\ \ \mbox{\bf{}proj}\ \ \aone\ \ \cdots\ \ \G_n\ \ \mbox{\bf{}proj}\ \ \an\\
\it \vdash&\it P(\G_{\rm1},\ldots ,\G_n)\ \ \mbox{\bf{}proj}\ \ \ti\makebox{\tt (}\aone\makebox{\tt ,...,}\an\makebox{\tt )}\ea}{\mu(\G_{\rm1},\ldots,\G_n).P(\G_{\rm1},\ldots,\G_n)\ \ \mbox{\bf{}proj}\ \ \ai}\espace\\
\it \ \ \ \ \ [\mbox{\rm{}where}\ \ai\makebox{\tt =}\ti\makebox{\tt (}\aone\makebox{\tt ,...,}\an\makebox{\tt )}]
\eeqs
}
For recursively-defined types the rules yield only those
projections that act on each recursive instance of a data structure in
the same way.  Thus ${\it{}FProj}_{\subtypespace{\makebox{\subtypesize\tt FunList}} }$ comprises $\BOT$ and
$\SPINE~\A$ for $\A$ ranging over $\BOT$ and $\ID$, where
\beqs
\it \SPINE\ \A\ =\ \mu\G.(\itt{NIL}\ \ID)\ \glb\ (\itt{CONS}\ (\A\ \x\ \G))\ ,
\eeqs
so $\SPINE\ \BOT$ specifies static spines and dynamic elements, and
$\SPINE\ \ID$ is the identity.  More generally, the abstract list
constructor is isomorphic to lifting.  Similarly,
${\it{}FProj}_{\subtypespace{\makebox{\subtypesize\tt FunTree}} }$
comprises $\BOT$ and $\LBR\ \A$ for $\A$ ranging over $\BOT$ and $\ID$, where
\beqs
\it \LBR\ \A\ =\ \mu\G.(\itt{LEAF}\ \A)\ \glb\ (\itt{BRANCH}\ (\G\ \x\ \G))\ ,
\eeqs
so $\LBR\ \A$ specifies static branches and leaves, and $\A$ of the
leaf nodes.  Again, the abstract tree constructor is isomorphic to
lifting.  This compares favourably with
BHA strictness analysis, for which the corresponding abstract
constructors are typically {\it double\/} lifting \cite{Wad87,Sew93}.

Value $\K$ is in ${\it{}FFor}_{\subtypespace\stt}$ if $\K~\bfabsf~\mbox{\tt T}$ can be
inferred from the following.

There is only one forward value at type $\tint$.
\beqs
\it \axm{()\ {}\ \bfabsf\ {}\ \tint}\ .
\eeqs
For products and sums,
\beqs
\it \infer{\K_{\rm1}\ {}\ \bfabsf\ {}\ \tone\ {\ }\ \cdots\ {\ }\ \K_n\ {}\ \bfabsf\ {}\ \tn}{(\K_{\rm1},\ldots,\K_n)\ {}\ \bfabsf\ {}\ \tprod}\ ,
\eeqs
%\espace\espace
\beqs
\it \infero{\K_{\rm1}\ {}\ \bfabsf\ {}\ \tone\ {\ }\ \cdots\ {\ }\ \K_n\ {}\ \bfabsf\ {}\ \tn}{(\K_{\rm1},\ldots,\K_n)\ {\ }\ \bfabsf\ {}\ \tsum}\ .
\eeqs
Function spaces consist of a set of step functions closed under lub.
\beqs
\it \infero{\ba{ll}\T_{\rm1}\in{\it{}FTran}_{\subtypespace\stone}\ {\ {\ }\ }\ \K_{\rm1}\ \bfabsf\ \tone\\
\it \T_{\rm2}\in{\it{}FTran}_{\subtypespace\sttwo}\ {\ {\ }\ }\ \K_{\rm2}\ \bfabsf\ \ttwo\ea}{\it{}step\ ((\T_{\rm1},\K_{\rm1}),(\T_{\rm2},\K_{\rm2}))\ {}\ \bfabsf\ {}\ \tfunp}\ ,
\eeqs
where
\beqs
\it step\ (v_{\rm1},v_{\rm2})\ x\ =\ v_{\rm2},&\it {\rm\ if}\ v_{\rm1}\we{}x\\
\it step\ (v_{\rm1},v_{\rm2})\ x\ =\ \bot,&\it {\rm\ otherwise}\ ,
\eeqs
and
\beqs
\it \infer{\K_{\rm1}\ {}\ \bfabsf\ {}\ \tfunp\ {\ {\ }\ }\ \K_{\rm2}\ {}\ \bfabsf\ {}\ \tfunp}{(\K_{\rm1}\lub\K_{\rm2})\ {}\ \bfabsf\ {}\ \tfunp}\ .
\eeqs
This gives the full space of monotonic functions on the abstract
domains.

For recursively-defined types, roughly speaking, we choose those
forward values that represent each component of the same type by the
same value.  Given type definitions $\tdefn$, which we will write
$\mbox{\tt A}_i\mbox{\tt =T}_i\mbox{\tt (A}_1\mbox{\tt ,...,A}_n\mbox{\tt )},\ \oin$, if by assuming
$\K_i~\bfabsf~\ai$ for $\oin$ we may deduce
$P_i(\K_1,\ldots,\K_n)~\bfabsf~\ti(\aone\ldots\an)$ for $\oin$, then
\beqs
\it \mu(\K_{\rm1},\ldots ,\K_n).(P_{\rm1}(\K_{\rm1},\ldots ,\K_n),\ldots ,P_n(\K_{\rm1},\ldots ,\K_n))
\eeqs
is a tuple $(\K_1,...,\K_n)$ of values such that $\K_i\ \bfabsf\ \mbox{\tt A}_i$
for $\oin$.

For all \mbox{\tt T} the set ${\it{}FProj}_{\subtypespace\stt}\times{\it{}FFor}_{\subtypespace\stt}$ is a 
finite lattice containing the top and bottom elements of $\Te{\Ps}{\mbox{\tt T}}$.

For \mbox{\tt T} not containing \mbox{\tt ->} the domain $\Tde{\Ps}{\mbox{\tt T}}$ is isomorphic
to $\one$, so ${\it{}FFor}_{\subtypespace\stint}$ is $\one$.  For \mbox{\tt Int}\ \mbox{\tt ->}\ \mbox{\tt Int}
we have
$\Tde{\Ps}{\mbox{\tt Int}\ \mbox{\tt ->}\ \mbox{\tt Int}}=(\proj{\De{}{\tint}}\x\one)\fto(\proj{\De{}{\tint}}\x\one)$,
so ${\it{}FFor}_{\subtypespace{\makebox{\subtypesize\tt Int->Int}} }$ is
$({\it{}FProj}_{\subtypespace\stint}\x\one)\fto({\it{}FProj}_{\subtypespace\stint}\x\one)$.  A
data structure of recursive type \mbox{\tt A=T(A)} may be thought of as some
(possibly infinite) number of elements of \mbox{\tt T(())}.  For example, the
value $\itt{cons}\ (f,\ \itt{cons}\ (g,\ \itt{nil}\ ()))$ in the
standard domain for \mbox{\tt FunList} decomposes into $\itt{cons}\ (f,())$,
$\itt{cons}\ (g,())$, and $\itt{nil}\ ()$.  The (implicit) abstraction
function maps such a data structure to the greatest lower bound of
these elements, giving a safe abstraction of the nonstandard values of
each element.  Thus
${\it{}FFor}_{\subtypespace{\makebox{\subtypesize\tt A=T(A)}} }={\it{}FFor}_{\subtypespace{\makebox{\subtypesize\tt T(())}} }$, so
${\it{}FFor}_{\subtypespace{\makebox{\subtypesize\tt FunList}} }\iso{\it{}FFor}_{\subtypespace{\makebox{\subtypesize\tt FunTree}} }\iso{\it{}FFor}_{\subtypespace{\makebox{\subtypesize\tt Int->Int}} }$.

\section{Examples of Analysis}

For all closed expressions \mbox{\tt e} the abstract value
$\Ee{\Ps}{\mbox{\tt e}}~[\,]$ of \mbox{\tt e} is of the form $(\ID,\K)$, showing
that closed expressions are always entirely static.  For expressions of
function type the abstract forward value $\K$ is a function from
abstract arguments of \mbox{\tt e} to abstract results.

First we consider functions on lists.
Let expression \mbox{\tt length} denote usual length function:
\if 0\theisrr
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ fix\ ({\char'134}length\ .\ {\char'134}xs\ .\ case\ xs\ of}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ nil\ u\ \ ->\ 0}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ cons\ p\ ->\ let\ (z,zs)\ =\ p\ in}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 1\ +\ length\ zs)\ .}\vspace{-0.0em}
\end{flushleft}
\else
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ fix\ ({\char'134}length\ .\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ {\char'134}xs\ .\ case\ xs\ of}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ nil\ u\ \ ->\ 0}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ cons\ p\ ->\ let\ (z,zs)\ =\ p\ in}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 1\ +\ length\ zs)\ .}\vspace{-0.0em}
\end{flushleft}
\fi
The abstract forward value of \mbox{\tt length} is of the form
$\L(\A,\K).(\T~\A,())$, where $\T$ maps $\SPINE~\BOT$ and $\SPINE~\ID$
to $\ID$, and $\BOT$ to $\BOT$.  This reveals that the result of
\mbox{\tt length} is independent of the values of list elements, and gives
a static result when the argument has a static spine.

Let \mbox{\tt append} stand for the expression denoting the usual function for
appending two lists:
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ fix\ ({\char'134}append\ .\ \ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ {\char'134}xs\ .\ {\char'134}ys\ .\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ case\ xs\ of}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ nil\ u\ \ ->\ ys}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ cons\ p\ ->\ let\ (z,zs)\ =\ p\ in}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ cons\ (z,\ append\ zs\ ys))\ .}\vspace{-0.0em}
\end{flushleft}
The abstract value of \mbox{\tt append} is
\beqs
\it (\ID,\ \L(\A_{\subtypespace{\makebox{\subtypesize\tt xs}} },\K_{\subtypespace{\makebox{\subtypesize\tt xs}} }).\\
\it (\ID,\ \L(\A_{\subtypespace{\makebox{\subtypesize\tt ys}} },\K_{\subtypespace{\makebox{\subtypesize\tt ys}} }).\\
\it (\A_{\subtypespace{\makebox{\subtypesize\tt xs}} }\glb\A_{\subtypespace{\makebox{\subtypesize\tt ys}} },\ \K_{\subtypespace{\makebox{\subtypesize\tt xs}} }\glb\K_{\subtypespace{\makebox{\subtypesize\tt ys}} })))\ .
\eeqs
This reveals that partial applications of \mbox{\tt append} are static up to WHNF, and the
abstract value of the result is the greatest lower bound of the two
arguments.  In general, the abstract value of a closed expression
of the form $\mbox{\tt {\char'134}x}_1\mbox{\tt .{\char'134}x}_2\ldots\mbox{\tt .e}$ will reveal that all partial
applications of the expression are static up to WHNF.

Let \mbox{\tt reverse1} stand for the expression denoting the naive reverse function:
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ fix\ ({\char'134}reverse1\ .\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ {\char'134}xs\ .\ case\ xs\ of}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ nil\ u\ ->\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ nil\ ()}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ cons\ p\ ->\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ let\ (z,zs)\ =\ p\ in}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ append\ (reverse\ zs)\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ (cons\ (z,\ nil\ ())))\ .}\vspace{-0.0em}
\end{flushleft}
The abstract forward value of \mbox{\tt reverse1} is the identity, so the abstraction of
the elements of a list doesn't change by reversing the list.

Let \mbox{\tt compose} stand for \mbox{\tt {\char'134}f.{\char'134}g.{\char'134}x.f\ (g\ x)}.  The abstract value of \mbox{\tt compose} is
\beqs
\it (\ID,\ \L(\A_{\subtypespace{\makebox{\subtypesize\tt f}} },\K_{\subtypespace{\makebox{\subtypesize\tt f}} }).\\
\it (\ID,\ \L(\A_{\subtypespace{\makebox{\subtypesize\tt g}} },\K_{\subtypespace{\makebox{\subtypesize\tt g}} }).\\
\it (\ID,\ \left\{\ba{lll}\bot,&\it \mbox{\rm\ if}\ \A_{\subtypespace{\makebox{\subtypesize\tt f}} }=\BOT\mbox{\rm\ or}\ \A_{\subtypespace{\makebox{\subtypesize\tt g}} }=\BOT\\
\it \K_{\subtypespace{\makebox{\subtypesize\tt f}} }\o\K_{\subtypespace{\makebox{\subtypesize\tt g}} },&\it \mbox{\rm\ otherwise}\ea\right\})))\ .
\eeqs
Thus, the
result of the application of the composition of two functions is
dynamic, and maps all values to dynamic values, if either function is
dynamic; otherwise, the result is given by the application of the
composition of the abstract forward values to the abstract argument.

Let \mbox{\tt listcomp} stand for the expression denoting the function
that composes lists of functions:
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ fix\ ({\char'134}listcomp\ .\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ {\char'134}fs\ .\ case\ fs\ of}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ nil\ u\ ->\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ {\char'134}x.x}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ cons\ p\ ->\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ let\ (g,gs)\ =\ p\ in\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ compose\ g\ (listcomp\ gs))\ .}\vspace{-0.0em}
\end{flushleft}
The abstract value of \mbox{\tt listcomp} is
\beqs
\it (\ID,\ \L(\A_{\subtypespace{\makebox{\subtypesize\tt fs}} },\K_{\subtypespace{\makebox{\subtypesize\tt fs}} }).\\
\it (\ID,\ \left\{\ba{lll}\bot,&\it \mbox{\rm\ if}\ \A_{\subtypespace{\makebox{\subtypesize\tt fs}} }\neq\SPINE\ \ID\\
\it \glb_{i\geq{\rm0}}(\K_{\subtypespace{\makebox{\subtypesize\tt fs}} })^i,&\it \mbox{\rm\ otherwise}\ea\right\}))\ .
\eeqs
Since the abstract values of lists contain no information about the
length of the lists of which they are abstractions, the abstract value
of the composition of list elements is the glb of the composition over
all lengths.

\commentout{
Next we consider functions on trees.
Let \mbox{\tt count} be short for
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ fix\ ({\char'134}count\ .\ {\char'134}t\ .\ case\ t\ of}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ leaf\ l\ \ \ ->\ 1}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ branch\ p\ ->\ let\ (t1,t2)\ =\ p\ in}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ count\ t1\ +\ count\ t2)\ ,}\vspace{-0.0em}
\end{flushleft}
denoting the function that counts leaves.  The abstract forward value
\mbox{\tt count} is of the form $\L(\A,\K).(\T~\A ())$, where $\T$ maps
$\LBR~\BOT$ and $\LBR~\ID$ to $\ID$, and $\BOT$ to $\BOT$.  This
reveals that the result of \mbox{\tt count} is independent of the values of leaf
elements, and maps trees with static branch and leaf structure, but
possibly dynamic values at the leaves, to static results.
}

Let \mbox{\tt flatten} stand for the expression denoting the function that flattens
trees into lists:
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ fix\ ({\char'134}flatten\ .\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ {\char'134}t\ .\ case\ t\ of}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ leaf\ l\ \ \ ->\ cons\ (l,\ nil\ ())}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ branch\ p\ ->\ let\ (t1,t2)\ =\ p\ in}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ append\ (flatten\ t1)\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ (flatten\ t2))\ .}\vspace{-0.0em}
\end{flushleft}
The abstract value of \mbox{\tt flatten} is
\beqs
\it (\ID,\ \L(\A_{\subtypespace{\makebox{\subtypesize\tt fs}} },\K_{\subtypespace{\makebox{\subtypesize\tt fs}} }).\\
\it (\T\ \A,\ \left\{\ba{lll}\bot,&\it \mbox{\rm\ if}\ \T\ \A_{\subtypespace{\makebox{\subtypesize\tt fs}} }\neq\LBR\ \ID\\
\it \K_{\subtypespace{\makebox{\subtypesize\tt fs}} },&\it \mbox{\rm\ otherwise}\ea\right\}))\ ,
\eeqs
where $\T$ maps $\BOT$ to $\BOT$, and $\LBR~\A$ to $\SPINE~\A$ for $\A$
ranging over $\ID$ and $\BOT$.

The function denoted by \mbox{\tt compose\ listcomp\ flatten} that composes
trees of functions has abstract value
\beqs
\it (\ID,\ \L(\A_{\subtypespace{\makebox{\subtypesize\tt fs}} },\K_{\subtypespace{\makebox{\subtypesize\tt fs}} }).\\
\it (\ID,\ \left\{\ba{lll}\bot,&\it \mbox{\rm\ if}\ \A_{\subtypespace{\makebox{\subtypesize\tt fs}} }\neq\LBR\ \ID\\
\it \glb_{i\geq{\rm0}}(\K_{\subtypespace{\makebox{\subtypesize\tt fs}} })^i,&\it \mbox{\rm\ otherwise}\ea\right\}))\ .
\eeqs
Similarly to the case for lists, the abstract values of trees contain
no information about the structure of the trees of which they are abstractions,
so the abstract value of the composition of the values of the leaves
is the glb of the composition over all tree structures.

\commentout{
Let \mbox{\tt treecomp} stand for the expression denote the function
that composes trees of functions:
\begin{flushleft}
\vspace{-0.0em}\makebox{\progsize\tt \ \ \ \ \ fix\ ({\char'134}treecomp\ .\ {\char'134}t\ .\ case\ fs\ of}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ leaf\ f\ \ \ ->\ f}\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ branch\ p\ ->\ let\ (t1,t2)\ =\ p\ in\ }\vspace{-0.0em}\\
\makebox{\progsize\tt \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ compose\ (listcomp\ t1)\ (listcomp\ t2))\ ,}\vspace{-0.0em}
\end{flushleft}
The abstract value of \mbox{\tt treecomp} is simply
\beqs
\it (\ID,\ \L(\A_{\subtypespace{\makebox{\subtypesize\tt fs}} },\K_{\subtypespace{\makebox{\subtypesize\tt fs}} }).(\ID,\ \left\{\ba{lll}\bot&\it \ \ \ \ \ \ \mbox{\rm\ if}\ \A_{\subtypespace{\makebox{\subtypesize\tt fs}} }=\BOT\\
\it \K_{\subtypespace{\makebox{\subtypesize\tt fs}} }&\it \ \ \ \ \ \ \mbox{\rm\ otherwise}\ \ea\ \right\}))\ .
\eeqs
That is ...}

\section{More on Abstract Domains}

The sizes of the abstract domains and the representations of the
abstract values can be considerably optimised.  The nonstandard
semantics of application---embodied by $\it apply^{\Ps}$---guarantees
that the abstract values $(\BOT,f)$ and $(\BOT,f')$ from
$\Te{\Ps}{\tfun}$, for all $f$ and $f'$, are effectively the same:
$\it apply^{\Ps}~(\BOT,f)=\bot$ for all $f$.  A practical analyser
would take advantage of this fact, identifying $(\BOT,f)$ over
all $f$.  More generally, for function types embedded
within data structures, e.g.\ $\mbox{\tt (}\tint\mbox{\tt ,}\tfun\mbox{\tt )}$, abstract values
$((\A,()),(\BOT,f))$ would be identified over all $f$.

In the following we restrict attention to {\it denotable\/} values:  at
each type \mbox{\tt T} those values that can be expressed as $\Ee{}{\mbox{\tt e}}~[]$ for
some \mbox{\tt e}.  For every value $v\in\Te{\Ss}{\mbox{\tt T}}$ there is a best
abstraction---a greatest value $(\A,\K)\in\Te{\Ps}{\mbox{\tt T}}$ such that
$\Re{}{\mbox{\tt T}}((\A,\K),v,v)$.  For types $\tone$ and $\ttwo$ not
containing \mbox{\tt ->} we have
$\Tde{\Ps}{\tfun}\iso(\proj{\De{}{\tone}}\x\one)\fto(\proj{\De{}{\ttwo}}\x\one)$,
and if $\K\in\Tde{\Ps}{\mbox{\tt T}}$ is greatest such that
$\Rde{}{\tfun}~(\K,f,f)$ for denotable $f$, then $\K$ maps $(\ID,())$
to $(\ID,())$ and distributes over $\glb$.  The subset of functions
from $\Tde{\Ps}{\tfun}$ that map $(\ID,())$ to $(\ID,())$ and
distribute over $\glb$ forms a complete lattice, hence attention may be
restricted to this subset.  Not only does this reduce the number of
abstract functions, it also reduces their representation:  such
functions are determined by their behaviour on the $\glb$-{\it basis\/}
of the lattice $\proj{\De{}{\tone}}$---the set $B$ of values such that
every element of $\proj{\De{}{\tone}}$ is the glb of some (possibly
empty) subset of $B$, and no element of $B$ can be expressed as the glb
of some subset not containing that element.  This optimisation can be
generalised to higher order:  $\Tde{\Ps}{\tfun}$ may be restricted to
functions that map $\top$ to $\top$ and distribute over $\glb$, for all
function types $\tfun$.  (These results follow from those shown in
\cite{Dav94}.)

\section{Related Work}

Consel \cite{Con90} describes a binding-time analysis for higher-order
untyped languages.  As in our analysis abstract values have two parts,
the first describing the static/dynamic properties of values, and the
second describing how (for function types) abstract arguments are
mapped to abstract results; there appears to be an implicit domain
factorisation similar to ours, based on implicit type information
collected by analysis semantics.  In this respect there are many
superficial similarities between the two techniques.  No formal
relation to the standard semantics is given, making a formal comparison
with our technique problematic.

\section{Conclusion}

We have successfully generalised Launchbury's monomorphic
projection-based binding-time analysis to higher-order, using
abstract domains smaller than those typically used in BHA
strictness analysis.

The next step would be to generalise to handle Hindley-Milner
polymorphism.  This has been done with good results at first order for
binding-time analysis \cite{Lau91a}, and for BHA strictness analysis
at higher order \cite{Bar91,Bar93}.  We anticipate that the combined
use of these theories will give a reasonably straightforward generalisation
to polymorphism.  In addition to making the analysis more widely applicable,
this should also greatly reduce the run-time cost of analysis.

Though not developed here, using the same approach it is
possible to give a strictness analysis technique, again closely related to
Hunt's PER-based strictness analysis technique \cite{Hun91}.

\section{Acknowledgements}

Thanks to the referees for their constructive comments on the first
version of this paper.

\begin{thebibliography}{123456}

\bibitem[Abr89]{Abr89}
S. Abramsky. 
``The lazy lambda calculus.'' 
{\it Research Topics in Functional Programming.}
David Turner, ed., Addison-Wesley 1989.

\bibitem[Abr90]{Abr90}
S. Abramsky. 
``Abstract interpretation, logical relations and Kan extensions.''
{\it Journal of Logic and Computation}, 1, 1990.

\bibitem[Bar91]{Bar91}
G. Baraki.
``A note on abstract interpretation of polymorphic functions.''
In \cite{Hug91}.

\bibitem[Bar93]{Bar93}
G. Baraki.
{\it Abstract Interpretation of Polymorphic Higher-Order Functions.}
Ph.D. thesis, Research report FP-1993-7, 
Department of Computing Science, University of Glasgow.

\bibitem[BEJ88]{BEJ88}
D. Bj{\odash}rner, A.P. Ershov, and N.D. Jones, eds.
{\it Partial Evaluation and Mixed Computation,
Proceedings IFIP TC2 Workshop, Gammel Avern{\ae}s}, Denmark, October 1987.
North-Holland, 1988.

\bibitem[Con90]{Con90}
C. Consel. 
``Binding Time Analysis for Higher Order Untyped Functional Languages.''
Proceedings of the 1990 ACM Conference on LISP and 
Functional Programming, pp264-272.

\bibitem[Dav93]{Dav93b}
K. Davis.
``Higher-order Binding-time Analysis.'' 
{\it Proceedings of the 1993 ACM on Partial Evaluation and
Semantics-Based Program Manipulation (PEPM '93)},
ACM Press, 1993.

\bibitem[Dav94]{Dav94}
K. Davis.\ \ 
{\it Projection-based Program Analysis.}
Thesis submitted for degree of Ph.D., Computing Science Department,
University of Glasgow, 1994.

\bibitem[Go92]{Go92}
C.K. Gomard.
``A self-applicable partial evaluator for the lambda calculus:  Correctness
and pragmatics.''
ACM TOPLAS, Vol 14, No.\ 2, April 1992.

\bibitem[HM94]{HM94}
F. Henglein and C. Mossin.
``Polymorphic binding-time analysis.''
European Symposium on Programming (ESOP '94).

\bibitem[Hug91]{Hug91}
J. Hughes, ed.
{\it Proceedings of the 1991 Conference on Functional Programming Languages
and Computer Architecture (FPCA '91)}, Cambridge, Sept 1991. LNCS 523,
Springer Verlag, 1991.

\bibitem[Hun91]{Hun91}
S. Hunt.
``PERs generalise projections for strictness analysis (extended abstract).''
{\it Proceedings of the 1990 Glasgow Workshop
on Functional Programming}.
Simon L. Peyton Jones {\it et al.}, eds.
Springer Workshops in Computing.
Springer-Verlag, 1991.

\bibitem[HS91]{HS91}
S. Hunt and D. Sands.
``Binding time analysis:  a new PERspective.''
{\it ACM Symposium on Partial Evaluation
and Semantics-Based Program Manipulation},
SIGPLAN Notices Vol.\ 26, No.\ 9, 1991.

\bibitem[Jen92]{Jen92}
T. Jensen.
{\it Abstract Interpretation in Logical Form}.
Ph.D. thesis, Report 93/11, Department of Computer Science,
University of Copenhagen, 1992.

\bibitem[Jon88]{Jon88}
N.D. Jones.
``Automatic program specialization: A re-examination from
basic principles.'' In \cite{BEJ88}.

\bibitem[Lau88]{Lau88}
J. Launchbury.
``Projections for specialisation.''
In \cite{BEJ88}.

\bibitem[Lau91a]{Lau91a}
J. Launchbury. 
{\it Projection Factorisations in Partial Evaluation.} PhD
Thesis, Glasgow University, Nov 89. Distinguished Dissertation in
Computer Science, Vol 1, CUP, 1991.

\bibitem[Lau91c]{Lau91c}
J. Launchbury.
``A strongly-typed, self-applicable partial evaluator.''
In \cite{Hug91}.

\bibitem[Mog89]{Mog89}
T. Mogensen.
``Binding-time analysis for polymorphically typed higher order languages.''
{\it International Joint Conference on Theory and Practice of Software
Development}. LNCS 352. Springer-Verlag 1989.

\bibitem[NN88]{NN88}
H.R. Nielson and F. Nielson.
``Automatic binding-time analysis for a typed $\L$-calculus.''
{\it Science of Computer Programming 10},
North Holland, 1988.  Also in POPL '88.

\bibitem[Sch88]{Sch88}
D.A. Schmidt.
``Static properties of partial reduction.''
In \cite{BEJ88}.

\bibitem[Sew93]{Sew93}
J. Seward.
``Polymorphic strictness analysis using frontiers.''
{\it Proceedings of the 1993 ACM on Partial Evaluation and
Semantics-Based Program Manipulation (PEPM '93)},
ACM Press, 1993.

\bibitem[Wad87]{Wad87}
P. Wadler.
``Strictness analysis on non-flat domains by abstract interpretation over finite domains. '' 
S. Abramsky, C. Hankin, eds.
{\it Abstract Interpretation of Declarative Languages.}
Ellis-Horwood, 1987.

\end{thebibliography}

\end{document}